---
title: "AT4BR exercise 6"
author:
  name: Weronika Oprzędek
format: html
editor: visual
---

Loading libraries

```{r}
library(tidyr)
library(dplyr)
library(readr)
library(rmarkdown)
```

## 1. Loading dataset

```{r}
data("iris")

paged_table(iris)
```

## 2. Transforming the wide table to a long table

I merge columns that contain petal or sepal size (length or width) into two columns and name them Trait with the category and Size with its value. This table is saved as iris_long.

```{r}
iris_long <- iris %>% 
  pivot_longer(cols=matches("Petal|Sepal"), names_to = "Trait", values_to = "Size")

iris_long
```

## 3.A) Mean length and width of the sepal and petal of each species

In the first step, I group the data by species and their traits. In the second step, for each group a mean value is counted and showed in a table.

```{r}
iris_long  %>% 
  group_by(Species, Trait) %>% 
  summarise(Mean_size = mean(Size))

```

## 3.B) Largest and widest, petal and sepal, of virginica and versicolor species

First, I filter the data to have only chosen species, then group them by their species and traits (petal and sepal, length and width) and in the end I search for a maximum value in each group and show the results in the table.

```{r}
iris_long  %>% 
  filter(Species %in% c("virginica", "versicolor")) %>% 
  group_by(Species, Trait) %>% 
  summarise(Max_size = max(Size))
```

## 3.C) The number of observations above the mean for each species, for sepal and petal, width and length

After grouping the data by species and traits and next calculating a mean for each group, I check for each value from the column Size if it's higher than the mean from it's group. It gives me data where in the place of each "Size" there's a TRUE or FALSE value. Then I sum up all the true values and display the final table.

```{r}
iris_long  %>% 
  group_by(Species, Trait) %>% 
  summarise(Above_the_mean = sum(Size > mean(Size)))
```

## 
